ms_figures

### 2018-02-20 man/ms_figures.Rmd
add changes to fig and upload new template

---
output:
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: template_ms.docx
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

@@

library("grid")
library("gridExtra")
library("patchwork")
```

@@

## Figure 2. Plot combinado con EVI profile (a) y anomalías (b)
")
```

### Figure 1

@@

### Figure 2

```{r}
iv <- read.csv(here("data/evi", "iv_composite.csv"), header = TRUE, sep = ',')

@@

```{r, fig.height=7}
a_profile_compara <- profile_compara + labs(title = "a")
b_plot_sa_clu <- plot_sa_clu + labs(title = "b")
a_profile_compara + b_plot_sa_clu + plot_layout(ncol=1)
```

**Figure 2.** Comparison of EVI profile for the reference period (*gray*) and during the 2005 (*green*) and 2012 (*blue*) drought events (a) (mean of all analyzed pixels). EVI standardized anomaly ($EVI~sa$) during the period 2000-2016 for northern and southern populations (b). Error bars show standard error.

### Figure 3

```{r}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

@@

```{r}
# EVI_drought 
res_evi_drought <-
  ggplot(mhuber_filter, aes_disturb) + 
  theme_res + theme_evi + theme(legend.position='top') + 
  ggtitle("a) EVI")

@@

```{r, fig.height=7}
res_evi_drought  + res_bai_drought + plot_layout(ncol=1)
```

**Figure 3.** Response of northern (*black*) and southern (*blue*) populations of *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) for the years 2005 and 2012. Different letters above bars indicate significant post hoc differences between groups (see material and methods).

### Figure 4
```{r}
theme_de <- list(
  scale_color_manual(values=c('black','red')), 
  scale_shape_manual(values=c(19, 15)),
  scale_fill_manual(values=c('black','red')),  
  scale_linetype_manual(values=c("solid", "solid")),
  labs(color = 'Drougth event', fill='Drougth event', linetype='Drougth event', shape='Drougth event')

@@

```{r}
res_evi_site  + res_bai_site + plot_layout(ncol=1)
```

**Figure 4.** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods).

### Figure 5

@@

cronos_bai <- cro_sites %>% ggplot(aes(x=year, y=bai_mean/100, colour=site)) + 
  theme_bw() + ylab('BAI') + 
  geom_ribbon(aes(ymin = (bai_mean - bai_se)/100,
                  ymax = (bai_mean + bai_se)/100,
                  fill=site), alpha=.2, colour=NA) +
  geom_line() + 
  geom_point(size=.8) + 
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(size=6),
        legend.position = c(0.1,.8),
        strip.background = element_blank()) + 
  scale_colour_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_fill_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  ylab(expression(paste("BAI (", cm^2, year^-1, ")", sep=''))) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10)) 
  # scale_y_continuous(sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  # geom_line(aes(x=year, y=n_trees+40, colour=site))

samp <- cro_sites %>% ggplot(aes(x=year, y=n_trees, colour=site)) + 
  theme_bw() + ylab('# trees') + 
  geom_line() + 
  theme(panel.grid = element_blank(),
        legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        plot.margin = margin(t = 5, r = 0, b = -5, l = 0, unit = "pt")) + 
  scale_colour_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_fill_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10))

  # theme(plot.margin = unit(c(1,5,-30,6),units="points"),
  #       axis.title.y = element_text(vjust =0.25))

plot_bai_s <- samp + cronos_bai + plot_layout(ncol=1, heights = c(0.1, 0.9))

plot_bai_s
```

**Figure 5.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA_Low; *pink*) and high-elevation (CA_High, *purple*) sites. Shading areas coorespond to standard error of the mean. Number of series are displayed in the upper plot.

# Figure 6

```{r}
cro_rwi <- read.csv(here("data/cronos_medias", "cronos_sites_rwi.csv"), header = TRUE, sep = ',')
year_cutoffs <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_cutoffs.csv"), header = TRUE, sep = ',')
year_sampledepth <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_sampledepth.csv"), header = TRUE, sep = ',')
```

@@

```{r}
label_site <- c("0_sj" = "Northern (SJ)",
                "1_caH" = "Southern High (CaH)",
                "2_caL" = "Southern Low (CaL)")

# 
plot_rwi_crono <- ggplot(cro_rwi, aes(x=year, y=res)) +
  theme_bw() + 
  theme(panel.grid = element_blank(), 
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face="bold", size=12), 
        axis.text = element_text(face="bold", size=10), 
        axis.title = element_text(face="bold", size=12)) + 
  geom_hline(yintercept = 1, colour = "grey") + 
  geom_line(size=.8) + scale_y_continuous(breaks = seq(0.4,1.8, 0.2)) +
  facet_wrap(~site_sorted, scales = "fixed", ncol = 1,
             labeller = as_labeller(label_site)) + 
  ylab("Ring width index") + xlab("Year") + 
  geom_vline(xintercept = c(1995, 2005, 2012), linetype = "dotted") + 
  geom_vline(aes(xintercept = year), data = year_cutoffs,  linetype = "dashed", colour='red') + 
  geom_vline(aes(xintercept = year), data = year_sampledepth,  linetype = "dashed", colour='blue') +
  scale_x_continuous(limits = c(1819, 2016), 
                     breaks= seq(1820, 2020, by=20))

plot_rwi_crono

**Figure 6.** Residual Tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the three most recent severe drought years (1995, 2005 and 2012).

# Figure 7
```{r}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',')

@@

**Figure 7.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data (precipitation (a), SPEI (b), maximun (c) and minimun (d) temperatures) for northern site (*green* bars), low-elevation southern site (*light blue* bars) and high-elevation shouthern (*dark blue* bars) site. Asteriks indicate significant ($P < 0.05$) correlation coefficients.

# Figure 8

@@

c <- gc_todos %>% 
  filter(site != 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("1_caH" = "Southern High (CaH)", "2_caL" = "Southern Low (CaL)"))) + 
    ylab(custom_ylab) + custom_corrected_gc

@@

**Figure 8.** Comparison of median growth change ($GC$) following [@Nowacki1997] for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods).

# Figure 9

@@

**Figure 9**. Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis).

>>>>

---
output:
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: template_ms_new_page.docx
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

@@

library("grid")
library("gridExtra")
library("patchwork")
library("ggcorrplot")
library("stringr")
library("pander")
library("lemon")
```

@@

## Figure 2. Plot de anomalias
- [X] Solamente Anomalias estandarizadas

## Figure 3. 
")
```

#####

@@

#####

```{r, eval=FALSE}
iv <- read.csv(here("data/evi", "iv_composite.csv"), header = TRUE, sep = ',')

@@

```{r, fig.height=7, eval=FALSE}
a_profile_compara <- profile_compara + labs(title = "a")
b_plot_sa_clu <- plot_sa_clu + labs(title = "b")
a_profile_compara + b_plot_sa_clu + plot_layout(ncol=1)
```

```{r, fig.height=5}
plot_sa_clu 
```

**Figure 2.** EVI standardized anomaly ($EVI~sa$) during the period 2000-2016 for northern and southern populations. Error bars show standard error.

##### 
```{r, eval=FALSE, echo=FALSE}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

@@

```{r, eval=FALSE, echo=FALSE}
# EVI_drought 
res_evi_drought <-
  ggplot(mhuber_filter, aes_disturb) + 
  theme_res + theme_evi + theme(legend.position='top') + 
  ggtitle("a) EVI")

@@

```{r, eval=FALSE, fig.height=7}
res_evi_drought  + res_bai_drought + plot_layout(ncol=1)
```

```{r, fig.height=7}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

aux <- mhuber_filter %>% mutate(vari = '0_EVI')
aux2 <- mhuberbai_filter %>% mutate(vari = '1_BAI')
m <- bind_rows(aux2, aux)

pd <- position_dodge(width = 0.6)

aes_disturb <- aes(x = disturb_year, y = M.Huber, 
                   color = site,  
                   fill = site, linetype=site,
                   shape = site, 
                   ymin = lower.ci, ymax = upper.ci)

variables  <- c("0_EVI" = "EVI", "1_BAI" = "BAI")

themeR <- list(
  geom_errorbar(width=.1, position = pd),
  geom_point(position = pd),
  theme_bw(),
  facet_wrap(~vari, labeller = labeller(vari = variables)),
  xlab(''), 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    strip.text = element_text(face="bold", size = 14),
    axis.title.y = element_text(face="bold", size = 14),
    plot.title = element_text(hjust = -0.05, size = 16, margin = margin(t=-5, b=1))),
  scale_shape_manual(values=c(19, 21, 17, 22, 15)),
  scale_color_manual(values=c('black', "blue", 'black','blue','blue')),
  scale_fill_manual(values=c('black', 'white', 'black','white','blue')),
  scale_linetype_manual(values=c("solid","solid","solid","solid", "solid")),
  scale_size_manual(values=c(2,2,2.5, 2.5, 2.5))
)

rt <- m %>% filter(var == "rt") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resistance") + 
  geom_text(aes(label=Letter, y = 1.1), position=pd, show.legend = FALSE)

rc <- m %>% filter(var == "rc") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Recovery")+
  geom_text(aes(label=Letter, y = 1.65), position=pd, show.legend = FALSE)

rs <- m %>% filter(var == "rs") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resilience") + xlab('Drought event') +
  geom_text(aes(label=Letter, y = 1.2), position=pd, show.legend = FALSE) 
```

```{r, fig.height=6}
grid_arrange_shared_legend(rt, rc, rs,ncol = 1, nrow = 3, position='top')
```

**Figure 3.** Response *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; left-plots) and tree radial growth (BAI; right-plots) for the years 2005 and 2012. For EVI we compared northern populations (black fill circle) with southern ones (blue empty circle). For BAI we compared northern population (San Juan, SJ; black triangle) with southerns populations: Cáñar-High (Ca-H; blue empty squares) and Cáñar-Low (Ca-L; blue fill sauares). Different letters above error bars indicate significant post hoc differences between groups (see material and methods).

```{r, eval=FALSE}
### Figure 4

theme_de <- list(
  scale_color_manual(values=c('black','red')), 
  scale_shape_manual(values=c(19, 15)),
  scale_fill_manual(values=c('black','red')),  
  scale_linetype_manual(values=c("solid", "solid")),
  labs(color = 'Drougth event', fill='Drougth event', linetype='Drougth event', shape='Drougth event')

@@

```{r, eval=FALSE}
res_evi_site  + res_bai_site + plot_layout(ncol=1)

**Figure 4.** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods). 
```

#####

@@

# cores 
cronos_bai <- cro_sites %>% filter(n_trees >5) %>% 
  ggplot(aes(x=year, y=bai_mean/100, colour=site)) + 
  theme_bw() + ylab('BAI') + 
  geom_ribbon(aes(ymin = (bai_mean - bai_se)/100,
                  ymax = (bai_mean + bai_se)/100,
                  fill=site), alpha=.2, colour=NA) +
  geom_line() + 
  geom_point(size=.8) + 
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(size=6),
        legend.position = c(0.1,.8),
        strip.background = element_blank()) + 
  scale_colour_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_fill_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  ylab(expression(paste("BAI (", cm^2, year^-1, ")", sep=''))) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10), limits=c(1815,2020)) +
  geom_smooth(data = subset(cro_sites, year > 1975 & site == 'CA_High'), 
              aes(x=year, y=bai_mean/100, colour=site), method = 'lm', se=FALSE) 
  # scale_y_continuous(sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  # geom_line(aes(x=year, y=n_trees+40, colour=site)) 
samp <- cro_sites %>% ggplot(aes(x=year, y=n_trees, colour=site)) + 
  theme_bw() + ylab('# trees') + 
  geom_line() + 
  theme(panel.grid = element_blank(),
        legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_text(size=6),
        # axis.text.x=element_blank(),
        # axis.ticks.x=element_blank(), 
        # plot.margin = margin(t = 5, r = 0, b = -5, l = 0, unit = "pt")
        ) + 
  scale_colour_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_fill_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10))  
  # theme(plot.margin = unit(c(1,5,-30,6),units="points"),
  #       axis.title.y = element_text(vjust =0.25))

plot_bai_s <- samp + cronos_bai + plot_layout(ncol=1, heights = c(0.15, 0.85))
plot_bai_s
```

```{r, eval=FALSE}

# Lm CA-High
df <- cro_sites %>% 
  filter(year >= 1975) %>% 
  filter(site == 'CA_High')

m1 <- lm(bai_mean ~ year, df)
summary(m1)
r2 <- format(summary(m1)$r.squared, digits = 3)
```

**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA_Low; *pink*) and high-elevation (CA_High, *purple*) sites. Shading areas coorespond to standard error of the mean. Number of series are displayed in the upper plot. We only show chronologies with # trees > 5.

#####

```{r}
cro_rwi <- read.csv(here("data/cronos_medias", "cronos_sites_rwi.csv"), header = TRUE, sep = ',')
year_cutoffs <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_cutoffs.csv"), header = TRUE, sep = ',')
year_sampledepth <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_sampledepth.csv"), header = TRUE, sep = ',')

# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% filter(d_duration > 2) %>% filter(maxyear > 1940 ) %>% as.data.frame() 
dys <- dys[,'maxyear'] 
```

@@

```{r}
label_site <- c("0_sj" = "Northern (SJ)",
                "1_caH" = "Southern (Ca-High)",
                "2_caL" = "Southern Low (Ca-Low)")

cro_rwi <- cro_rwi %>% 
  mutate(siteF = case_when(
    site_sorted == "0_sj" ~ "Northern (SJ)",
    site_sorted == "1_caH" ~ "Southern (Ca-High)",
    site_sorted == "2_caL" ~ "Southern (Ca-Low)"))

# 
plot_rwi_crono <- ggplot(cro_rwi, aes(x=year, y=res)) +
  theme_bw() + 
  theme(panel.grid = element_blank(), 
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face="bold", size=12), 
        # strip.text = element_blank(), 
        axis.text = element_text(face="bold", size=8), 
        axis.title = element_text(face="bold", size=12)) + 
  geom_hline(yintercept = 1, colour = "grey") + 
  geom_line(size=.8) + scale_y_continuous(breaks = seq(0.4,1.8, 0.2)) +
  facet_wrap(~site_sorted, scales = "free_x", ncol = 1,
             labeller = as_labeller(label_site)) + 
  ylab("Ring width index") + xlab("Year") + 
  geom_vline(aes(xintercept = year), data = year_cutoffs,  linetype = "dashed", colour='red') + 
  # geom_vline(aes(xintercept = year), data = year_sampledepth,  linetype = "dashed", colour='blue') +
  geom_vline(xintercept = c(1995, 2005, 2012), linetype = "dotted") + 
  # geom_vline(xintercept = dys, linetype = "dotted") + 
  scale_x_continuous(limits = c(1819, 2016), 
                     breaks= seq(1820, 2020, by=20))

plot_rwi_crono  
```

**Figure 5.** Residual Tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the three of most recent severe drought years (1995, 2005 and 2012).

##### 
```{r}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',')

@@

**Figure 6.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data (precipitation (a), SPEI (b), maximun (c) and minimun (d) temperatures) for northern site (*green* bars), low-elevation southern site (*light blue* bars) and high-elevation shouthern (*dark blue* bars) site. Asteriks indicate significant ($P < 0.05$) correlation coefficients.

#####

@@

c <- gc_todos %>% 
  filter(site != 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("1_caH" = "Southern (CA-High)", "2_caL" = "Southern(CA-Low)"))) + 
    ylab(custom_ylab) + custom_corrected_gc

@@

**Figure 7.** Comparison of median growth change ($GC$) following [@Nowacki1997] for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods).

#####

@@

**Figure 8**. Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis).

# Supplementary

##### 
**Appendix S1** Acumulated monthly precipitation during the hydrological year 2004-2005 (blue line) and 2011-2012 (red line). The boxplot representing the average from 1940-2015 period. Data from meteorological station Granada, Base Aérea (National Spanish Meteorological Services, AEMET).

```{r, fig.height=5}
ltdf <- read.table(file=here("/data_raw/meteo", "long_term_data_rediam_data.txt"), header=T, sep=';', dec=',')

source(here("/script/R/", "hydro_year.R"))

ltdf <- ltdf %>% 
  dplyr::select(cod = INDICATIVO, VARIABLE, FECHA, VALOR) %>% 
  dplyr::mutate(fecha = as.Date(FECHA, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

# We used only the 5514 station (the most long term)
# Filter by station and prec 
# Filter from 1950 
st5514 <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1950) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR)

st5514_pro <- st5514 %>% 
  group_by(hyear_f, hmonth) %>% 
  dplyr::summarise(pre = sum(pre)) %>%
  dplyr::mutate(csum = cumsum(pre)) %>% as.data.frame()

st5514_pro  %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012'))) %>% 
  ggplot(aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot(aes(x=as.factor(hmonth))) +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Precipitation (mm) acumulated') + xlab('') + 
  ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 6, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 6, colour = "red") + 
  annotate("text", label = "mean 1950-2015", x = 3, y = 600, size = 6, colour = "black")
```

#####

```{r, echo=FALSE}
# Get file names
myfiles <- list.files(path = here("/data_raw/spei_all/"), pattern = "\\.csv$")

mydf <- data.frame()

# Loop to read files and prepare data
for (j in myfiles){ 
  aux <- read.csv(file=paste0(here("/data_raw/spei_all/"), j), header = TRUE, sep = ',')

  # Remove csv and get name 
  name_aux <- str_replace(j, ".csv", "")

  # Get lat long
  latlong <- str_replace(name_aux, "SPEI_", "")

  mylat <- as.numeric(str_split_fixed(latlong, "_", 2))[1]
  mylong <- as.numeric(str_split_fixed(latlong, "_", 2))[2]

  aux$lat <- mylat
  aux$long <- mylong

  # Split time 
  # Get year 
  aux$year <- as.numeric(str_extract(aux$DATA, "[0-9]+"))

  # Get month 
  aux$month <- str_extract(aux$DATA, "[aA-zZ]+")

  # Variable month in number 
  aux$months <- match(aux$month, month.abb)

  # assign(name_aux, aux)

  mydf <- rbind(mydf, aux)
}

spei <- mydf %>% select(SPEI_12, year, months) %>% 
  group_by(year,months) %>% 
  summarise(spei6 = mean(SPEI_12),
            spei6_sd = sd(SPEI_12)) %>% 
  mutate(date = lubridate::ymd(paste(year, months, "01")))

spei <- spei %>% 
  mutate(sign = case_when(
    spei6 > 0 ~ "pos",
    TRUE ~ "neg"))
```

```{r, fig.height=5}
limits <- as.Date(c("1949-01-01", "2015-12-01"))
spei %>% 
  ggplot(aes(x=date, y=spei6, fill=sign)) +  
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) +
  theme_bw()  + 
  labs(title= 'Drought evolution for Sierra Nevada',
    x= '',
    y='SPEI') +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none") +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y", limits=limits) 
```

**Appendix S2**. Drought severity in the Sierra Nevada for the 1950-2016 period based on the Standardised Precipitation-Evapotranspiration Index (SPEI). We used data from Global SPEI database. We obtanied the SPEI data for a 12 month scale and for all 0.5º grid cells covering Sierra Nevada.

```{r, echo=FALSE, eval=FALSE}
**Figure S2**. Spatial distribution of the EVI-annual trends ($\tau$) for the period 2000-2016. Negative values (*red* values) indicate negative trends and positive values (*green* values) indicate positive trends.

![](../analysis/evi/compute_MannKendall_evimean_files/figure-markdown_github/rater_map_tau_EVI-1.png){width=150%}
```

##### 
```{r}
prec_files <- list.files(here::here('data_raw/meteo/prec_subclima/'), pattern="*.csv")

dfpre <- c()

for (i in prec_files){

  auxdf <- read.csv(here::here('data_raw/meteo/prec_subclima', i), header = TRUE, sep=";")
  auxdf$code <- gsub(x=i, pattern = ".csv", replacement = "")

  dfpre <- rbind(dfpre, auxdf)
}

dfpre$date <- as.Date(dfpre$date, format = "%d/%m/%Y")

dfpre <- dfpre %>% 
  dplyr::mutate(fecha = as.Date(date, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

aux <- dfpre %>% 
  group_by(hyear_f, code) %>% 
  dplyr::mutate(csum = cumsum(prec)) %>% 
  filter(hmonth == 12) %>% 
  dplyr::select(-month, -prec, -date, -fecha) %>% as.data.frame()

presn <- aux %>% group_by(hyear) %>% 
     dplyr::summarise(
     csum_mean = mean(csum, na.rm = TRUE),
     csum_sd = sd(csum, na.rm = TRUE), 
     csum_se = csum_sd/sqrt(length(csum_sd)), 
     n = n())

meandf <- mean(presn$csum_mean)
sddf <- sd(presn$csum_mean)

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() +
  scale_x_continuous(breaks=seq(1950,2020, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) +
  geom_text_repel(
    data = subset(presn, csum_mean < (meandf - 1*sddf)),
    aes(label = hyear)) +
  theme(panel.grid.minor = element_blank()) +
  ylab('Cumulative Precipitation (hydrological year)') + xlab('year')
prec_hydrol_sn
```

**Appendix S3.** Temporal evolution of cumulative precipitation (hydrological year) during the period 1950-2017 for 28 meteorological stations (from National Spanish Meteorological Services, AEMET) distributed around the Sierra Nevada area. Points represent mean and errorbars standard error. *Black* line indicates mean for all period. *Red* lines represent -1 (*dotted* line) and -2 (*dashed* line) standard deviation. *Blue* lines represent +1 (*dotted* line) and +2 (*dashed* line) standard deviation. We indicate all years with average values below -1SD.

##### 
```{r}
# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% filter(d_duration > 2) %>% as.data.frame()

pander(dys)
```

**Appendix S4.** Tabla de drought severas.

##### 
**Appendix S5.** Percentage of pixels showing browning, greenning or no-changes during the 2005 and 2012 droguht events according to EVI standardized anomalies.

```{r}
df <- read.csv(file=here::here('/out/anomalies/evi' ,'percen_browing_greening.csv'), header=TRUE)

df <- df %>% add_row(clu_pop2 = 'Northern slope',
                             type = 'greening',
                             count_clu = 0, per = 0.00, y = 2005)

per_pixels_green <- df %>% 
  ggplot(aes(x=type, y=per, fill=as.factor(y))) + 
  geom_bar(stat='identity', position = "dodge") + 
  geom_text(aes(label = paste0(per, " %")), position=position_dodge(width=0.9), hjust=-.1, size=3, angle = 90) + 
  facet_wrap(~clu_pop2) + 
  theme_bw() + xlab('') + ylab('Frequency of pixels (%)') +
  theme(strip.background = element_rect(colour = "black", fill = "white")
        ) + labs(fill = 'Drought event') + scale_fill_manual(values = c("#DAA520", "#104E8B")) +
  scale_y_continuous(expand=c(.1,0))

per_pixels_green
```

#####

**Appendix S6.** Reilience vs. droughts

##### 
**Appendix S7.** Correlation between indices of resilience (Rt, resistance; Rc, recovery; Rs, Resilience) using periods of several lengths (2, 3 and 4 years after a drought). Top plots (a, b and c) showing the resilience indices of greenness (EVI) to drought; and bottom plots (d, e, f) the resilience indices of tree-growth (BAI) to drought. ws2, ws3 and ws4 indicate periods of 2, 3 and 4 years after a drought.

```{r}

corBai <- read.csv(file=here("/out/correla_resilience/", "correla_window_size.csv"), header = TRUE)
corEvi <- read.csv(file=here("/out/correla_resilience/", "correla_window_size_evi.csv"), header = TRUE)
corRwi <- read.csv(file=here("/out/correla_resilience/", "correla_window_size_rwi.csv"), header = TRUE)

generaMatrix <- function(cordf, variable){ 
  m <- cordf %>% filter(var == variable) %>% 
    dplyr::select(-var, -window_size) %>% 
    mutate(w1 = c(2, 2, 3), 
           w2 = c(3,4,4)) 
  mf <- m %>% 
    dplyr::select(r2, w2, w1) %>% rename(w1 = w2, w2 = w1)

  df <- m %>% rbind(mf) %>% 
    add_row(r2 = 1, w1=2, w2=2) %>% 
    add_row(r2 = 1, w1=3, w2=3) %>% 
    add_row(r2 = 1, w1=4, w2=4)

    df <- df %>% 
    mutate(w1 = case_when(
      w1 == 2 ~ 'ws2',
      w1 == 3 ~ 'ws3',
      w1 == 4 ~ 'ws4'),
      w2 = case_when(
      w2 == 2 ~ 'ws2',
      w2 == 3 ~ 'ws3',
      w2 == 4 ~ 'ws4'))

      cor_df <- as.data.frame.matrix(xtabs(r2 ~., df))

return(cor_df)
}

cor_rt <- generaMatrix(corBai, 'rt')
cor_rc <- generaMatrix(corBai, 'rc')
cor_rs <- generaMatrix(corBai, 'rs')

cor_rte <- generaMatrix(corEvi, 'rt')
cor_rce <- generaMatrix(corEvi, 'rc')
cor_rse <- generaMatrix(corEvi, 'rs')

g_rt <- ggcorrplot(cor_rt, type='upper', lab=TRUE, title='BAI Rt', show.legend = FALSE)
g_rc <- ggcorrplot(cor_rc, type='upper', lab=TRUE, title='BAI Rc', show.legend = FALSE)
g_rs <- ggcorrplot(cor_rs, type='upper', lab=TRUE, title='BAI Rs', show.legend = FALSE)
g_rte <- ggcorrplot(cor_rte, type='upper', lab=TRUE, title='EVI Rt', show.legend = FALSE)
g_rce <- ggcorrplot(cor_rce, type='upper', lab=TRUE, title='EVI Rc', show.legend = FALSE)
g_rse <- ggcorrplot(cor_rse, type='upper', lab=TRUE, title='EVI Rs', show.legend = TRUE)

(g_rte + g_rce + g_rse) / (g_rt + g_rc + g_rs) + plot_layout(nrow=2) + plot_annotation(tag_levels = c('a'))
```

##### 
**Appendix S8.** Tabla S1

##### 
**Appendix S9.** Tabla S2
